---
title: "RT-Glo Cytation 2020-10-08 analysis"
author: "Darren Tyson"
date: "03/10/2021"
output: html_notebook
---

#### Data from Cytation experiment performed 2020-12-14
Combined luminescence and imaging reads of several SCLC and melanoma cell lines treated with various drugs and concentrations.  

NOTE: Some contamination (bacterial? fungal?) was visible in some cell culture vessels prior to experiment (per Clayton Wandishin).  


```{r Setup}
library(diprate)

closestTime <- function (mytime, timevec, direction = "", out = "pos")
{
    sapply(mytime, function(mt) 
    {
        difft <- switch(direction,
                        before = -difftime(timevec, mt),
                        after = -difftime(mt, timevec),
                        abs(difftime(mt, timevec))
                        )
        if(direction %in% c("before","after"))
        {
            mindiff = ifelse(length(difft[difft>=0]) == 0, NA, min(difft[difft>=0]))
            if(is.na(mindiff))
            {
                r <- NA
            } else {
                r <- as.vector(switch(out,
                                      time = timevec[which(difft == min(difft[difft>=0]) )],
                                      pos = which(difft == min(difft[difft>=0])),
                                      amt = min(difft[difft>=0])
                                      )
                               )
            }
        } else {
            r <- switch(out,
                        time = as.character(timevec[which(difft == min(difft))]),
                        pos = which(difft == min(difft)),
                        amt = min(difft)
                        )
        }
        return(r)
    })
}

getDateTime <- function(dname)
{
    o <- dname[grepl("[0-9]{6}_*[0-9]{6}",dname)]
    if(all(grepl("[0-9]{12}",o)))
    {
        mytime <- strptime(o, '%y%m%d%H%M%S')
    } else {
        mytime <- paste(sapply(strsplit(o,'_'), '[[', 1),
            sapply(strsplit(o,'_'), '[[', 2), sep='_')
        mytime <- strptime(mytime, '%y%m%d_%H%M%S')
    }
    names(mytime) <- NULL
    return(mytime)
}
```

#### Data assembly
Data were not completely compiled into a single dataframe for analysis. Must assemble prior to analysis.

```{r}
TOPDIR <- "../data/20201216_Lum_Image_Run/"
# Momentum log path
ml_path <- file.path(TOPDIR,'MOMENTUM_LOGS_FOR_V004831B_20201214_100msRFP_10msGFP_20201216_153632_0.txt')
# Momentum log
ml <- read.csv(ml_path, header=FALSE, skip=5, as.is=TRUE, sep='\t')
# Index of rows with reads
read_i <- which(ml$V6=='Read')
# Keep only relevant columns
ml <- ml[read_i,c('V1','V6','V7','V12','V13','V14')]
# Rename columns
colnames(ml) <- c('instrument','action','id','start.time','end.time','barcode')
rownames(ml) <- NULL


lum <- read.csv(file.path(TOPDIR,'CombinedLuminescenceTOTAL.csv'), as.is=TRUE, row.names=1)
# Cell counts = cc
cc <- read.csv(file.path(TOPDIR,'Cellcounts_20201214.csv'), as.is=TRUE)
cc <- cc[order(cc$plate_id,cc$well),]

# standardize colnames
colnames(cc) <- gsub("_",".",colnames(cc))
```

#### Data munging
Luminescence data already has most annotation associated with it, including plate barcode (ID), cell line, drug, drug conc, time and well. The times were parsed from the filenames of the raw luminescence data (see [../data/20201216_Lum_Image_Run/MultiPlateMultiFileMergeLum20201216_special.py])  

Momentum log should have times matching (or very similar to) these values. Must make sure they're in the same format. Use the `closestTime` function to obtain the position in the 

The cell counts were obtained from two different positions within each well. These values should be summed and same for the counts of nuclei that are positive in the second channel (the dead-cell stain; `ch2_pos`).

```{r Time matching}
# Cell count image file path =  ccifp; from task arguments sent to py-seg image processing pipeline
ccifp <- read.csv(file.path(TOPDIR,'TaskArgs_20201214.csv'))$nuc_im_path
# Find times for each unique plate
ccifp <- gsub("/mnt/monica/quaranta2/Cytation/2020-12-14/images/","",ccifp)
ccifp <- unique(sapply(strsplit(ccifp, "/"), "[[", 1))

# get plate IDs from number after "Experiment"; should match Momentum log
pid <- as.integer(sapply(strsplit(ccifp,"Experiment"), "[[", 2))
names(pid) <- getDateTime(ccifp)

cc$image.time <- names(pid)[match(cc$plate.id,pid)]

# standardize lum colnames
colnames(lum) <- gsub("_",".",tolower(colnames(lum)))
if(!any(grepl("drug1",colnames(lum)))) colnames(lum) <- gsub("drug","drug1",colnames(lum))
colnames(lum)[colnames(lum)=="tothour.lum"] <- "time"
colnames(lum)[colnames(lum)=="datetime.lum"] <- "lum.time"

# ulfst = unique lum file save times
ulfst <- unique(lum$lum.time)

# plate name (barcode from Momentum log)
# pn <- ml[closestTime(unique(cc$image.time),ml[grepl('RFP',ml$id),'start.time.rfmt']),'barcode']
# plate name (from lum)
pn <- lum[closestTime(unique(cc$image.time),unique(lum$lum.time),direction="before"),'plate.name']
names(pn) <- unique(cc$image.time)

# add plate.name (barcode) to cc
cc$plate.name <- pn[match(cc$image.time,names(pn))]


# must sum cell counts and ch2_pos per uid2 (unique for each time point) 
cc$uid <- paste(cc$plate.id,cc$well,sep="_")
cc$uid2 <- paste(cc$uid,cc$plate.id,sep="_")
cc_temp <- cc[!duplicated(cc$uid2),]
cc_temp$cell.count <- sapply(unique(cc$uid2), function(x) sum(cc[cc$uid2==x,'cell.count']))
cc_temp$ch2.pos <- sapply(unique(cc$uid2), function(x) sum(cc[cc$uid2==x,'ch2.pos']))

# time lookup table: cc image times as names, lum file save times as values
time_lut <- closestTime(unique(cc_temp$image.time),ulfst,direction="before",out="time")

# Add luminescence read times; note that the first luminescence read is missing
cc_temp$lum.time <- time_lut[cc_temp$image.time]

# remove unneeded columns from cc_temp
cc_temp <- cc_temp[,c("plate.name","image.time","lum.time","well","cell.count","ch2.pos")]
# first plate did not get luminescence read; must pull info from Momentum logs or remove
cc_temp <- cc_temp[!is.na(cc_temp$plate.name),]
cc_temp$uid <- paste(cc_temp$plate.name,cc_temp$well,cc_temp$lum.time, sep="_")

lum$uid <- paste(lum$plate.name,lum$well,lum$lum.time, sep="_")




# ml$start.time.rfmt <- strptime(ml$start.time, format="%m/%d/%Y %r")
# ml$end.time.rfmt <- strptime(ml$end.time, format="%m/%d/%Y %r")
# ml$image.time <- unique(cc$image.time)[closestTime(ml$start.time.rfmt,unique(cc$image.time))]


```

## NOTE: plate names and times do not match between imaging and lum
Try to use momentum log to reconcile


### Stop here


#### Not yet incorporated with above
DO NOT RUN

```{r fig.height=5, fig.width=5}
cont <- d[d$drug1.conc==0,]
#do.call(plotGC, getGCargs(cont[cont$well %in% c('B11','C11','D11'),], 
#                          dat.col = c("time", "cell.count", "well")))
invisible(do.call(plotGC, getGCargs(cont, 
                          dat.col = c("time", "cell.count", "well"))))

```

# Sum cell counts across wells in same condition 
To facilitate visualization due to the low cell numbers visible within each field of view. 
```{r}
d$ucond <- paste(d$drug1,d$drug1.conc,d$time,sep='_')
a <- d[order(d$ucond),]
a_counts <- sapply(unique(a$ucond), function(x) sum(a[a$ucond==x,'cell.count']))
a_counts_ch2 <- sapply(unique(a$ucond), function(x) sum(a[a$ucond==x,'ch2.pos']))
a <- a[!duplicated(a$ucond),]
a$cell.count <- a_counts
a$ch2.pos <- a_counts_ch2
a <- a[order(a$well,a$time),]
```

### Examine control wells
```{r fig.height=5, fig.width=5}
cont <- a[a$drug1.conc==0,]
cont <- cont[order(cont$well,cont$time),]

invisible(do.call(plotGC, getGCargs(cont, 
                          dat.col = c("time", "cell.count", "well"))))
```

```{r fig.height=5, fig.width=5}
invisible(do.call(plotGC, append(getGCargs(a[a$drug1=='TAK-901',], 
                          dat.col = c("time", "cell.count", "drug1.conc")),
                          list(main='HEK293 + TAK-901', ylab='Population doublings'))))
```
#### Luminescence reads
Read in data, reformat and replace time with image time for direct comparisons
```{r}
lum <- read.csv("../data/HEK293lum.csv", as.is=TRUE)
lum_time <- lum[,1]
lum2 <- as.matrix(lum[,-1])
wells <- fixWellName(colnames(lum)[-1])
lum <- data.frame(time=rep(unique(d$time)), 
                       lum=as.vector(lum2), 
                       well=rep(wells, each=length(unique(d$time))))
lum$uid <- paste(match(lum$time, unique(lum$time)), lum$well, sep="_")
lum <- addMapInfo(lum,"../data/201008_185701_Experiment1_platemap.tsv")
```

```{r}
lum$ucond <- paste(lum$drug1,lum$drug1.conc,lum$time,sep='_')
a2 <- lum[order(lum$ucond),]
a2_lum <- sapply(unique(a2$ucond), function(x) sum(a2[a2$ucond==x,'lum']))
a2 <- a2[!duplicated(a2$ucond),]
a2$lum <- a2_lum
a2 <- a2[order(a2$well,a2$time),]
```

#### Combine all data
All data should be aligned by row such that can simply add luminescence reads to cell count data.
```{r}
if(all(a$uid==a2$uid)) 
{
    a$lum <- a2$lum
} else {
    message('Cell count data and luminescence data not aligned!')
}
```

```{r}
plot(a$cell.count, a$lum)
```
#### Only control wells exhibiting clear exponential growth
By visual inspection of cell count growth curves, combined wells B11, C11 and D11 (saved into B11) demonstrated exponential growth after ~60 h. Use
```{r fig.height=4, fig.width=4}
a2 <- a[a$well=="B11",]
plot(lum ~ cell.count, data=a2, 
     xlab="Cell count", ylab="Luminescence", main="HEK293 control wells")
```
Correlation of luminescence to cell counts in control wells
```{r}
round(cor(a2$cell.count,a2$lum),4)
```

### Save combined data file
Only save if file does not already exist.
```{r}
dfn <- "../data/2020-10-08_Cytation_data_sum_by_condition.csv"
if(!file.exists(dfn)) write.csv(a, file=dfn, row.names=FALSE)
```
Also save data with all wells included
```{r}
d$lum <- lum$lum
cfn <- "../data/2020-10-08_Cytation_data_counts+lum.csv"
if(!file.exists(cfn)) write.csv(d, file=dfn, row.names=FALSE)
```

