---
title: "RT-Glo Cytation 2020-10-08 analysis"
author: "Darren Tyson"
date: "03/10/2021"
output: html_notebook
---

#### Data from Cytation experiment performed 2020-12-14
Combined luminescence and imaging reads of several SCLC and melanoma cell lines treated with various drugs and concentrations.  

NOTE: Some contamination (bacterial? fungal?) was visible in some cell culture vessels prior to experiment (per Clayton Wandishin).  


```{r Setup}
library(diprate)

closestTime <- function (mytime, timevec, direction = "", out = "pos")
{
    sapply(mytime, function(mt) 
    {
        difft <- switch(direction,
                        before = -difftime(timevec, mt),
                        after = -difftime(mt, timevec),
                        abs(difftime(mt, timevec))
                        )
        if(direction %in% c("before","after"))
        {
            mindiff = ifelse(length(difft[difft>=0]) == 0, NA, min(difft[difft>=0]))
            if(is.na(mindiff))
            {
                r <- NA
            } else {
                r <- as.vector(switch(out,
                                      time = timevec[which(difft == min(difft[difft>=0]) )],
                                      pos = which(difft == min(difft[difft>=0])),
                                      amt = min(difft[difft>=0])
                                      )
                               )
            }
        } else {
            r <- switch(out,
                        time = timevec[which(difft == min(difft))],
                        pos = which(difft == min(difft)),
                        amt = min(difft)
                        )
        }
        return(r)
    })
}

```

#### Data assembly
Data were not completely compiled into a single dataframe for analysis. Must assemble prior to analysis.

```{r}
TOPDIR <- "../data/20201216_Lum_Image_Run/"
ml_path <- file.path(TOPDIR,'MOMENTUM_LOGS_FOR_V004831B_20201214_100msRFP_10msGFP_20201216_153632_0.txt')
ml <- read.csv(ml_path, header=FALSE, skip=5, as.is=TRUE, sep='\t')
read_i <- which(ml$V6=='Read')
ml <- ml[read_i,c('V1','V6','V12','V13','V14')]
colnames(ml) <- c('instrument','action','start.time','end.time','barcode')
rownames(ml) <- NULL

lum <- read.csv(file.path(TOPDIR,'CombinedLuminescenceTOTAL.csv'), as.is=TRUE, row.names=1)
cc <- read.csv(file.path(TOPDIR,'Cellcounts_20201214.csv'), as.is=TRUE)
cc <- cc[order(cc$plate_id,cc$well),]
cc$uid <- paste0(cc$plate_id)
```

#### Data munging
Luminescence data already has most annotation associated with it, including plate barcode (ID), cell line, drug, drug conc, time and well. The times were parsed from the filenames of the raw luminescence data (see [../data/20201216_Lum_Image_Run/MultiPlateMultiFileMergeLum20201216_special.py])  

Momentum log should have times matching (or very similar to) these values. Must make sure they're in the same format. Use the `closestTime` function to obtain the position in the 

The cell counts were obtained from two different positions within each well. These values should be summed and same for the counts of nuclei that are positive in the second channel (the dead-cell stain; `ch2_pos`).

```{r Time matching}
ml$acq.time <- strptime(ml$end.time, format="%m/%d/%Y %r")

lum$file.save.time <- strptime(lum$Datetime_Lum, format="%Y-%m-%d %H:%M:%S")
# ufst = unique file save times
ufst <- unique(lum$file.save.time)

# find closest time in ufst to each ml$acq.time
ml$match.time <- ufst[closestTime(ml$acq.time, ufst, direction="before")]

```


```{r fig.height=5, fig.width=5}
cont <- d[d$drug1.conc==0,]
#do.call(plotGC, getGCargs(cont[cont$well %in% c('B11','C11','D11'),], 
#                          dat.col = c("time", "cell.count", "well")))
invisible(do.call(plotGC, getGCargs(cont, 
                          dat.col = c("time", "cell.count", "well"))))

```

# Summe cell counts across wells in same condition 
To facilitate visualization due to the low cell numbers visible within each field of view. 
```{r}
d$ucond <- paste(d$drug1,d$drug1.conc,d$time,sep='_')
a <- d[order(d$ucond),]
a_counts <- sapply(unique(a$ucond), function(x) sum(a[a$ucond==x,'cell.count']))
a_counts_ch2 <- sapply(unique(a$ucond), function(x) sum(a[a$ucond==x,'ch2.pos']))
a <- a[!duplicated(a$ucond),]
a$cell.count <- a_counts
a$ch2.pos <- a_counts_ch2
a <- a[order(a$well,a$time),]
```

### Examine control wells
```{r fig.height=5, fig.width=5}
cont <- a[a$drug1.conc==0,]
cont <- cont[order(cont$well,cont$time),]

invisible(do.call(plotGC, getGCargs(cont, 
                          dat.col = c("time", "cell.count", "well"))))
```

```{r fig.height=5, fig.width=5}
invisible(do.call(plotGC, append(getGCargs(a[a$drug1=='TAK-901',], 
                          dat.col = c("time", "cell.count", "drug1.conc")),
                          list(main='HEK293 + TAK-901', ylab='Population doublings'))))
```
#### Luminescence reads
Read in data, reformat and replace time with image time for direct comparisons
```{r}
lum <- read.csv("../data/HEK293lum.csv", as.is=TRUE)
lum_time <- lum[,1]
lum2 <- as.matrix(lum[,-1])
wells <- fixWellName(colnames(lum)[-1])
lum <- data.frame(time=rep(unique(d$time)), 
                       lum=as.vector(lum2), 
                       well=rep(wells, each=length(unique(d$time))))
lum$uid <- paste(match(lum$time, unique(lum$time)), lum$well, sep="_")
lum <- addMapInfo(lum,"../data/201008_185701_Experiment1_platemap.tsv")
```

```{r}
lum$ucond <- paste(lum$drug1,lum$drug1.conc,lum$time,sep='_')
a2 <- lum[order(lum$ucond),]
a2_lum <- sapply(unique(a2$ucond), function(x) sum(a2[a2$ucond==x,'lum']))
a2 <- a2[!duplicated(a2$ucond),]
a2$lum <- a2_lum
a2 <- a2[order(a2$well,a2$time),]
```

#### Combine all data
All data should be aligned by row such that can simply add luminescence reads to cell count data.
```{r}
if(all(a$uid==a2$uid)) 
{
    a$lum <- a2$lum
} else {
    message('Cell count data and luminescence data not aligned!')
}
```

```{r}
plot(a$cell.count, a$lum)
```
#### Only control wells exhibiting clear exponential growth
By visual inspection of cell count growth curves, combined wells B11, C11 and D11 (saved into B11) demonstrated exponential growth after ~60 h. Use
```{r fig.height=4, fig.width=4}
a2 <- a[a$well=="B11",]
plot(lum ~ cell.count, data=a2, 
     xlab="Cell count", ylab="Luminescence", main="HEK293 control wells")
```
Correlation of luminescence to cell counts in control wells
```{r}
round(cor(a2$cell.count,a2$lum),4)
```

### Save combined data file
Only save if file does not already exist.
```{r}
dfn <- "../data/2020-10-08_Cytation_data_sum_by_condition.csv"
if(!file.exists(dfn)) write.csv(a, file=dfn, row.names=FALSE)
```
Also save data with all wells included
```{r}
d$lum <- lum$lum
cfn <- "../data/2020-10-08_Cytation_data_counts+lum.csv"
if(!file.exists(cfn)) write.csv(d, file=dfn, row.names=FALSE)
```

