---
title: "RT-Glo Cytation 2020-10-08 analysis"
author: "Darren Tyson"
date: "10/14/2020"
output: html_notebook
---

```{r Setup}
library(diprate)
```

Define loaction of files
```{r}
SEGDIR = file.path('/mnt/darren/quaranta2/Cytation/2020-10-08/Segmentation')
```

### Assemble cell count data
Need to pull all cellcounts.csv files together. 

```{r}
cc <- assemPlateData(SEGDIR, toFile=FALSE)
```
#### Filename structure
Example filename: `B10_04_1_1_RFP_001.tif`  

* `B10` = well  
* `04` = unknown  
* `1` = channel number (`1` or `2` in these data)  
* `1` = position number (`1` or `2` in these data)  
* `RFP` = channel name (`RFP` or `GFP` in these data)  
* `001` = time point index (only `001` in these data; actual time point index in enclosing directory (2 up) `Experiment[0-9]{1,2}`)  
* `tif` = image file format (only `tif` in these data)  


```{r}
d <- cc
```

```{r}
d$pos <- sapply(strsplit(d$file_name, "_"), "[[", 4)
d$uid <- paste(d$plate_id,d$well,sep='_')
ccw <- sapply(unique(d$uid), function(id) sum(d[d$uid==id,'cell_count']))
ch2w <- sapply(unique(d$uid), function(id) sum(d[d$uid==id,'ch2_pos']))
d <- d[d$pos==1,]
d$cell.count <- ccw
d$ch2.pos <- ch2w
d <- d[,c('plate_id','well','uid','cell.count','ch2.pos')]
```

Add other relevant columns and platemap info.
```{r}
d$expt.date <- "2020-10-08"
d <- addMapInfo(d,"../data/201008_185701_Experiment1_platemap.tsv")
```

Must get image acquisition times from directory names

First, make function to extract date and time from directory name.
```{r}
getDateTime <- function(dname)
{
    o <- dname[sapply(dname, function(d) grepl("[0-9]{6}_[0-9]{6}",d))]
    mytime <- paste(sapply(strsplit(o,'_'), '[[', 1), 
                    sapply(strsplit(o,'_'), '[[', 2), sep='_')
    mytime <- strptime(mytime, '%y%m%d_%H%M%S')
    names(mytime) <- NULL
    return(mytime)
}
```
Assemble directory names for each Experiment number

```{r}
im_time <- basename(list.dirs(dirname(SEGDIR), recursive=FALSE))
im_time <- im_time[!grepl('Segment',im_time)]
names(im_time) <- gsub('Experiment','',sapply(strsplit(im_time, '_'), '[[', 3))
im_time <- getDateTime(im_time)
```

Add image time to data
```{r}
d$image.time <- im_time[d$plate_id]
tx.time <- min(im_time) - (3600*6)
d$time <- signif(as.numeric(difftime(d$image.time,tx.time, units='hours')),4)
d <- d[order(d$well,d$time),]
rownames(d) <- NULL
```

```{r fig.height=5, fig.width=5}
cont <- d[d$drug1.conc==0,]
#do.call(plotGC, getGCargs(cont[cont$well %in% c('B11','C11','D11'),], 
#                          dat.col = c("time", "cell.count", "well")))
invisible(do.call(plotGC, getGCargs(cont, 
                          dat.col = c("time", "cell.count", "well"))))

```

# Summe cell counts across wells in same condition 
To facilitate visualization due to the low cell numbers visible within each field of view. 
```{r}
d$ucond <- paste(d$drug1,d$drug1.conc,d$time,sep='_')
a <- d[order(d$ucond),]
a_counts <- sapply(unique(a$ucond), function(x) sum(a[a$ucond==x,'cell.count']))
a_counts_ch2 <- sapply(unique(a$ucond), function(x) sum(a[a$ucond==x,'ch2.pos']))
a <- a[!duplicated(a$ucond),]
a$cell.count <- a_counts
a$ch2.pos <- a_counts_ch2
a <- a[order(a$well,a$time),]
```

### Examine control wells
```{r fig.height=5, fig.width=5}
cont <- a[a$drug1.conc==0,]
cont <- cont[order(cont$well,cont$time),]

invisible(do.call(plotGC, getGCargs(cont, 
                          dat.col = c("time", "cell.count", "well"))))
```

```{r fig.height=5, fig.width=5}
invisible(do.call(plotGC, append(getGCargs(a[a$drug1=='TAK-901',], 
                          dat.col = c("time", "cell.count", "drug1.conc")),
                          list(main='HEK293 + TAK-901', ylab='Population doublings'))))
```
#### Luminescence reads
Read in data, reformat and replace time with image time for direct comparisons
```{r}
lum <- read.csv("../data/HEK293lum.csv", as.is=TRUE)
lum_time <- lum[,1]
lum2 <- as.matrix(lum[,-1])
wells <- fixWellName(colnames(lum)[-1])
lum <- data.frame(time=rep(unique(d$time)), 
                       lum=as.vector(lum2), 
                       well=rep(wells, each=length(unique(d$time))))
lum$uid <- paste(match(lum$time, unique(lum$time)), lum$well, sep="_")
lum <- addMapInfo(lum,"../data/201008_185701_Experiment1_platemap.tsv")
```

```{r}
lum$ucond <- paste(lum$drug1,lum$drug1.conc,lum$time,sep='_')
a2 <- lum[order(lum$ucond),]
a2_lum <- sapply(unique(a2$ucond), function(x) sum(a2[a2$ucond==x,'lum']))
a2 <- a2[!duplicated(a2$ucond),]
a2$lum <- a2_lum
a2 <- a2[order(a2$well,a2$time),]
```

#### Combine all data
All data should be aligned by row such that can simply add luminescence reads to cell count data.
```{r}
if(all(a$uid==a2$uid)) 
{
    a$lum <- a2$lum
} else {
    message('Cell count data and luminescence data not aligned!')
}
```

```{r}
plot(a$cell.count, a$lum)
```
#### Only control wells exhibiting clear exponential growth
By visual inspection of cell count growth curves, combined wells B11, C11 and D11 (saved into B11) demonstrated exponential growth after ~60 h. Use
```{r fig.height=4, fig.width=4}
a2 <- a[a$well=="B11",]
plot(lum ~ cell.count, data=a2, 
     xlab="Cell count", ylab="Luminescence", main="HEK293 control wells")
```
Correlation of luminescence to cell counts in control wells
```{r}
round(cor(a2$cell.count,a2$lum),4)
```

### Save combined data file
Only save if file does not already exist.
```{r}
dfn <- "../data/2020-10-08_Cytation_data_sum_by_condition.csv"
if(!file.exists(dfn)) write.csv(a, file=dfn, row.names=FALSE)
```
Also save data with all wells included
```{r}
d$lum <- lum$lum
cfn <- "../data/2020-10-08_Cytation_data_counts+lum.csv"
if(!file.exists(cfn)) write.csv(d, file=dfn, row.names=FALSE)
```

